---
title: "SQL Exercises on Seattle and Portland Flights data"
author: "Dr. Chester Ismay"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(gradethis)
library(DBI)
library(RSQLite)
library(tidyverse)

# Copy data over into this folder
# file.copy(from = "../flights2023/pnwflights23/data/airlines.rda", 
#           to = "data/airlines.rda")
# file.copy(from = "../flights2023/pnwflights23/data/airports.rda",
#           to = "data/airports.rda")
# file.copy(from = "../flights2023/pnwflights23/data/flights.rda",
#           to = "data/flights.rda")
# file.copy(from = "../flights2023/pnwflights23/data/planes.rda",
#           to = "data/planes.rda")
# file.copy(from = "../flights2023/pnwflights23/data/weather.rda",
#           to = "data/weather.rda")

# Read data in from RDA files
load("data/airlines.rda")
load("data/airports.rda")
load("data/flights.rda")
load("data/planes.rda")
load("data/weather.rda")

## Create an ephemeral in-memory RSQLite database
con <- dbConnect(SQLite(), ":memory:")

# Write the new data to the database
dbWriteTable(con, "airlines", airlines, overwrite = TRUE)
dbWriteTable(con, "airports", airports, overwrite = TRUE)
dbWriteTable(con, "flights", flights, overwrite = TRUE)
dbWriteTable(con, "planes", planes, overwrite = TRUE)
dbWriteTable(con, "weather", weather, overwrite = TRUE)

knitr::opts_chunk$set(connection = "con", 
                      max.print = 20)
```

```{r setup2, echo=FALSE, out.width='100%', fig.align='center'}
tutorial_options(exercise.cap = "Query", 
                 exercise.eval = FALSE,
                 exercise.reveal_solution = TRUE)

library(DiagrammeR)
# https://github.com/bergant/datamodelr
library(datamodelr)

dm_f <- dm_from_data_frames(airlines, airports, flights, planes, weather)

graph <- dm_create_graph(dm_f, rankdir = "BT", col_attr = c("column", "type"))

dm_render_graph(graph)
```

# Selection Techniques

## Selecting columns/fields

1. Select all airline names from the `airlines` table.

```{sql query1, exercise=TRUE}

```

```{sql query1-solution}
SELECT name
  FROM airlines;
```

```{r query1-check}
grade_this({
  # Execute the solution query to get the correct result
  .solution <- DBI::dbGetQuery(con, .solution_code)

  # Execute the student's query
  .result <- DBI::dbGetQuery(con, .user_code)

  # Expected column
  cols_expected <- c("name")
  
  # Check for missing columns
  for (column in cols_expected) {
    if (!column %in% names(.result)) {
      fail(paste("Your query is missing the", column, "column."))
    }
  }
  
  # Check for extra columns
  extra_cols <- setdiff(names(.result), cols_expected)
  if (length(extra_cols)) {
    extra_cols <- knitr::combine_words(extra_cols, and = " or ", before = "`")
    fail(paste("Your query should not include the columns", extra_cols, "."))
  }
  
  # Check if the number of rows matches
  if (nrow(.result) != nrow(.solution)) {
    fail("The number of rows returned by your query does not match the expected result.")
  }

  # Check if the values match (assuming order doesn't matter)
  if (!all(sort(.result[[1]]) == sort(.solution[[1]]))) {
    fail("The values returned by your query do not match the expected result.")
  }
  
  # If all checks pass
  pass("Your query appears to be correct.")
})

```

## Aliasing

## Unique entries

## View creation

## Counting

# Filtering Techniques

## Filtering rows/records

## Filtering text

# Aggregating Techniques

## Numerical summaries

## Categorical summaries

## Rounding summaries

# Sorting and Grouping Techniques

## Sorting

## Grouping

# Transforming Techniques

# Joining Techniques

## INNER JOIN

## LEFT JOIN

## RIGHT JOIN

## Anti-Join
